A Computational Account of Default Mode Function\\ by Reinforcement Learning Theory

### 2017-03-26 paper/paper.tex

\section{Reinforcement Learning (RL) as a computational model for DMN function}
We now argue that the DMN has biological substrates (gradient
computation, backprop, experience replay, etc.) for all the
major components of an efficient \textit{Reinforcement Learning} algorithm, and all these
components integrate in a plausible way.
Reinforcement Learning (RL) is learning to achieve a goal by interacting with an external
environment. In a general RL framework, an \textit{agent}
interacts with an  environment in a trial-by-error fashion by taking
actions / decisions $a$, which trigger changes in the
\textit{state} of the environment
$s \rightarrow s'$, accompanied by a real-valued \textit{reward}
$r = r(s, a)$ (or regret!, since it can be negative) collected by the
agent. In this view, the environmental is partially controlled by
the action of the agent and the reward should be thought
of as an instantaneous satisfaction accompanying the execution of
an action, and may be delayed, only becoming substantial when accumulated over an extended period of time.
%% Here the subscript $t$ denotes time, and can be discrete or
%% continuous.
The environment is generally taken as stochastic,
is partially observed, in the sense that only part of the current state is observed by
the agent. This unpredictability
% (in a first-order sense)
is realistic in a model which sets out
to explain DMN function. Indeed, it is known that, the more the external world is predictable,
the more mental activity becomes detached from the actual sensorium
\citep{antrobus1966studies, pope1978regulation}.
Conversely, the more the currently executed task is unknown and unpracticed,
the lower the DMN activity, i.e the tendency for stimulus-independent thoughts
\citep{filler1973daydreaming, teasdale1995stimulus}.

These ``offline'' processes may however contribute to optimized control of the organism.
A DMN framework incorporating reinforcement learning
can naturally embed human behavior
into the tension between exploitative action with immediate gains and
explorative action with longer-term reward schedules.
The implication of the DMN in a diversity of human cognitive processes
can be parsimoneously recast as prediction error minimization
based on pervasive probabilistic simulations,
thus maximizing reward outcome at various time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo sampling
\citep{tenenbaum2011grow}.

>>>>

\section{Reinforcement Learning (RL) as a computational model for DMN function}
We now argue that the DMN has biological substrates (gradient
computation, backprop, experience replay, etc.) for all the
major components of an efficient \textit{Reinforcement Learning} algorithm, and all these
components integrate in a plausible way.
Reinforcement Learning (RL) is learning to achieve a goal by interacting with an external
environment. In a general RL framework, an \textit{agent}
interacts with an  environment in a trial-by-error fashion by taking
actions / decisions $a$, which trigger changes in the
\textit{state} of the environment
$s \rightarrow s'$, accompanied by a real-valued \textit{reward}
$r = r(s, a)$ (or regret!, since it can be negative) collected by the
agent. In this view, the environmental is partially controlled by
the action of the agent and the reward should be thought
of as an instantaneous satisfaction accompanying the execution of
an action, and may be delayed, only becoming substantial when accumulated over an extended period of time.
%% Here the subscript $t$ denotes time, and can be discrete or
%% continuous.
The environment is generally taken as stochastic,
is \textit{partially observed},
in the sense that only part of the current state is observed by
the agent. This unpredictability
% (in a first-order sense)
is realistic in a model which sets out
to explain DMN function. Indeed, it is known that, the more the external world is predictable,
the more mental activity becomes detached from the actual sensorium
\citep{antrobus1966studies, pope1978regulation}.
Conversely, the more the currently executed task is unknown and unpracticed,
the lower the DMN activity, i.e the tendency for stimulus-independent thoughts
\citep{filler1973daydreaming, teasdale1995stimulus}.
%
These ``offline'' processes may however contribute to optimized control of the organism.
A DMN framework incorporating reinforcement learning
can naturally embed human behavior
into the tension between exploitative action with immediate gains and
explorative action with longer-term reward schedules.
The implication of the DMN in a diversity of human cognitive processes
can be parsimoneously recast as prediction error minimization
based on pervasive probabilistic simulations,
thus maximizing reward outcome at various time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo sampling
\citep{tenenbaum2011grow}.
